---
title: The Genomic Signal of Host-Parasite Coevolution in Continuous Space
author:
  - name: Bob Week
    affil: 1
    orcid: '0000-0002-1099-3857'
  - name: Gideon Bradburd
    affil: 1
affiliation:
  - num: 1
    address: Department of Integrative Biology, Michigan State University
column_numbers: 2
logoright_name: SLiM_1024.jpg
logoleft_name: bradburd.jpg
primary_colour: "#1E3659"
author_textsize: "80px"
affiliation_textsize: "70px"
body_textsize: "60px"
output: 
  posterdown::posterdown_html:
    self_contained: false
bibliography: packages.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

Coevolution between hosts and parasites can maintain spatial correlations of allele frequencies at causal loci, a form of interspecific linkage disequilibrium (iLD). However, host-parasite systems that are distributed continuously in space may yield spurious iLD as a consequence of isolation-by-distance within each species. Here we compare distributions of iLD for neutral and causal loci to understand when coevolution may be detected using iLD.

# Methods

- Individual-based simulations performed using multispecies SLiM (see poster [281W](https://abs.genetics-gsa.org/pages/peqg22/Reports/ViewProgram/281))
- Causal loci accumulate following infinite-sites model
- Additive quantitative traits
- Gaussian distributed effect sizes
- Host-parasite trait-matching model (see Figure X)
- Intraspecific spatial competition
- Gaussian dispersal in two-dimensional geography

# Results

```{r mytable, out.width='80%'}
knitr::kable(iris[1:10, 1:4], caption = 'Table caption.',align = 'c',"html")
```

# Conclusion

# References
