---
title: "Likelihood of Heirarchical Matern Gaussian Random Field"
header-includes:
  - \usepackage{amsmath}
  - \usepackage{dutchcal}
  - \usepackage{mathrsfs}
  - \usepackage{csquotes}
  - \usepackage{textcomp}
  - \usepackage[T3,T1]{fontenc}
  - \DeclareSymbolFont{tipa}{T3}{cmr}{m}{n}
  - \DeclareMathAccent{\invbreve}{\mathalpha}{tipa}{16}
output: pdf_document
editor_options: 
  chunk_output_type: console
---

We model trait values of individuals as independent normally distributed variables around a mean $\bar z$ with variance $v$. In turn the mean $\bar z$ is a function of location $\pmb x=(x_1,x_2)^\top$ and follows a stationary Gaussian random field with spatially homogeneous mean $\tilde z$ and Matern covariance $C(d)$ with collocated variance $V=C(0)$, characteristic length $\xi$ and smoothness parameter $\nu=1$. Hence, observed individual trait values $z_1$ and $z_2$ at locations $\pmb x_1,\pmb x_2$ separated at a distance $d$ will be bivariate normally distributed with the following statistical properties:

$$\mathbb Ez_1=\mathbb Ez_2=\tilde z, \ \mathbb Vz_1=\mathbb Vz_2=v+V, \mathbb C(z_1,z_2)=C(d),$$

$$\mathbb E(z_i|\bar z)=\bar z(\pmb x_i), \ \mathbb V(z_i|\bar z)=v, \ \mathbb C(z_1,z_2|\bar z)=0.$$

Since we only get a single sample of this random field, the likelihood function is just

$$L(\tilde z,v,V,\xi|\pmb z)=\frac{1}{\sqrt{(2\pi)^n|\Sigma|}}\exp\left(-\frac{1}{2}(\pmb z-\pmb{\tilde z})\Sigma^{-1}(\pmb z-\pmb{\tilde z})^\top\right),$$

where $\pmb z=(z_1,\dots,z_n)^\top$ is the vector of observed individual trait values (here there are $n$ individuals) observed at locations $\pmb x_1,\dots,\pmb x_n$, $\pmb{\tilde z}$ is the vector of expected trait values $(\tilde z,\dots,\tilde z)^\top$ and $\Sigma$ is a covariance matrix with $v+V$ on the diagonal and $C(\|\pmb x_i-\pmb x_j\|)$ on the $i,j$th off-diagonal entry.

To maximize likelihood, we first compute gradients of the log-likelihood $\ell=\ln L$ with respect to model parameters. We then solve for the parameter values when those gradients are zero:

$$\frac{\partial\ell}{\partial\tilde z}=-\frac{1}{2}\frac{\partial}{\partial\tilde z}\left[(\pmb z-\pmb{\tilde z})\Sigma^{-1}(\pmb z-\pmb{\tilde z})^\top\right]=-\frac{1}{2}\frac{\partial}{\partial\tilde z}\sum_{ij}(\sigma^{-1})_{ij}(z_i-\tilde z)(z_j-\tilde z)$$

$$=\sum_{ij}(\sigma^{-1})_{ij}\Big(\tfrac{z_i+z_j}{2}-\tilde z\Big)=\sum_{ij}\tfrac{z_i+z_j}{2}(\sigma^{-1})_{ij}-\tilde z\sum_{ij}(\sigma^{-1})_{ij},$$

$$\implies \hat{\tilde z}=\frac{\sum_{ij}\tfrac{z_i+z_j}{2}(\sigma^{-1})_{ij}}{\sum_{ij}(\sigma^{-1})_{ij}}.$$

Unfortunately, the rest is analytically intractable. So we proceed via numerical optimization to maximize likelihood... in julia... then import the results below.

```{r, echo=F}
mle = read.csv("likelihood/mle.csv")
input = read.csv("likelihood/input.csv")

mlmeans.v = c()
mlmeans.V = c()
mlmeans.xi = c()
for(i in input$id){
  mlmeans.v = c( mlmeans.v, mean(mle$v[which(mle$id==i & mle$v!=Inf)]) )
  mlmeans.V = c( mlmeans.V, mean(mle$V[which(mle$id==i & mle$V!=Inf)]) )
  mlmeans.xi = c( mlmeans.xi, mean(mle$xi[which(mle$id==i & mle$xi!=Inf)]) )
}

cols = rep("purple",max(input$id))
cols[which((input$v/input$V)>1)] = "red"
cols[which((input$v/input$V)==1)] = "green"

cols.xi = rep("purple",max(input$id))
cols.xi[which(min(input$xi)==input$xi)] = "red"
cols.xi[which(max(input$xi)==input$xi)] = "green"
```

```{r, echo=F, fig.height=3}
par(mfrow=c(1,3))

plot(input$v,mlmeans.v, log="xy", bg=cols, pch=21, cex=1.5, xlab="Local phenotypic variance, v",ylab="MLE")
abline(0,1)

plot(input$V,mlmeans.V, log="xy", bg=cols, pch=21, cex=1.5, xlab="Collocated mean trait variance, V", ylab="")
abline(0,1)

plot(input$xi,mlmeans.xi, log="xy", bg=cols, pch=21, cex=1.5, xlab="Characteristic scale, xi", ylab="")
abline(0,1)
text(x=2.5,y=1e+98, labels=c("V=v"), col="black", cex=1.5)
text(x=1,y=1e+98, labels=c("*"), col="green",cex=4)
text(x=2.5,y=1e+85, labels=c("V<v"), col="black", cex=1.5)
text(x=1,y=1e+85, labels=c("*"), col="red",cex=4)
text(x=2.5,y=1e+70, labels=c("V>v"), col="black", cex=1.5)
text(x=1,y=1e+70, labels=c("*"), col="purple",cex=4)
```

```{r, echo=F, fig.height=3}
par(mfrow=c(1,3))

plot(input$v,mlmeans.v, log="xy", bg=cols.xi, pch=21, cex=1.5, xlab="Local phenotypic variance, v",ylab="MLE")
abline(0,1)

plot(input$V,mlmeans.V, log="xy", bg=cols.xi, pch=21, cex=1.5, xlab="Collocated mean trait variance, V", ylab="")
abline(0,1)

plot(input$xi,mlmeans.xi, log="xy", bg=cols.xi, pch=21, cex=1.5, xlab="Characteristic scale, xi", ylab="")
abline(0,1)
text(x=2.5,y=1e+98, labels=c("xi=10"), col="black", cex=1.5)
text(x=0.8,y=1e+98, labels=c("*"), col="green",cex=4)
text(x=2.5,y=1e+85, labels=c("xi=0.1"), col="black", cex=1.5)
text(x=0.8,y=1e+85, labels=c("*"), col="red",cex=4)
text(x=2.5,y=1e+70, labels=c("xi=1.0"), col="black", cex=1.5)
text(x=0.8,y=1e+70, labels=c("*"), col="purple",cex=4)
```

